---
title: "Intro to R Shiny"
author: "Kim Ky"
date: "11/29/2017"
output: html_document
runtime: shiny
---

## What is Shiny?

`shiny` is "an R package that makes it easy to build interactive web apps straight from R." [Click here](https://shiny.rstudio.com/) for more information.


## Install R Shiny

If you do not have `shiny` package installed on your machine, use command below to install. 

```{r install, eval=FALSE}
install.packages("shiny")
```

Once you have `shiny` installed, you can load the package using the `library` function.

```{r, eval=FALSE}
library(shiny)
```

In following examples, we will use `iris` data set that is available from MASS package.
```{r}
data("iris") # load iris data set
head(iris) # show first several rows of iris
```

## Setting up

A shiny app needs `ui` and `server`. They can be in the same file or two separate files that are named **ui.R** and **server.R**. We will start with defining `ui` and `server` in the same file. 

```{r}
ui <- fluidPage(
    titlePanel("This is an empty Shiny app.")
  )

server <- function(input, output, session) {}
```

Once you define `ui` and `server` for our app, you can run it using the `shinyApp` command as below.

```{r}
shinyApp(
  ui = ui,
  server = server
)
```

### Layout

Use `sidebarLayout` to create a layout with a sidebar and main areas. Usually, sidebar panel contains all the inputs (where users interact) and main panel contains the output based on the inputs.

```{r}
ui <- fluidPage(
  titlePanel("Side bar layout!"),
    sidebarLayout(
      sidebarPanel(p("This is the sidebar panel")),
      mainPanel(p("This is the main panel"))
    )
  )

shinyApp(ui = ui, server = server)
```

### Inputs

Let's make our Shiny app a little bit more interesting! 

```{r}
ui <- fluidPage(
  titlePanel("My Shiny App"),
    sidebarLayout(
      sidebarPanel(
        selectInput(inputId = 'x', label = "X Variable", choices = names(iris)),
        radioButtons(inputId = 'y', label = "Y Variable", choices = names(iris), inline = FALSE)
      ),
      mainPanel(p("This is the main panel"))
    )
  )
shinyApp(ui = ui, server = server)
```

### Output

Now we can add output based on selected inputs. 

```{r}
library(ggplot2)
ui <- fluidPage(
    titlePanel("My Shiny App"),
  sidebarLayout(
      sidebarPanel(
        selectInput(inputId = 'x', label = "X Variable", choices = names(iris)),
        radioButtons(inputId = 'y', label = "Y Variable", choices = names(iris), inline = FALSE)
      ),
      mainPanel(
        plotOutput(outputId = 'plot')
      )
    )
  )
server = function(input, output, session) {
  output$plot <- renderPlot({
    ggplot() + 
      geom_point(aes(x = iris[, input$x], y = iris[, input$y], size = 1), color = "#6f22b6") + 
      theme_bw() + 
      theme(legend.position = 'none') +
      labs(x = paste(input$x), y = paste(input$y), title = paste0("Plotting ", input$x, " by ", input$y))
  })
}
shinyApp(ui = ui, server = server)
```

### Observe vs Reactive

The main difference between `observe` and `reactive` is that `observe` does not return any values while `reactive` does. These are used to update output based on inputs "reactively".

```{r}
ui <- fluidPage(
    titlePanel("My Shiny App"),
  sidebarLayout(
      sidebarPanel(
        selectInput(inputId = 'x', label = "X Variable", choices = names(iris)),
        radioButtons(inputId = 'y', label = "Y Variable", choices = names(iris), inline = FALSE)
      ),
      mainPanel(
        plotOutput(outputId = 'plot')
      )
    )
  )
server = function(input, output, session) {
  observe({
    updateRadioButtons(session, 'y', choices = names(iris)[!names(iris) %in% input$x])
  })
  
  dat <- reactive({
    return(data.frame(x = iris[, input$x], y = iris[, input$y]))
  })
  
  output$plot <- renderPlot({
    ggplot(dat()) + 
      geom_point(aes(x = x, y = y, size = 1), color = "#6f22b6") + 
      theme_bw() + 
      theme(legend.position = 'none') +
      labs(x = paste(input$x), y = paste(input$y), title = paste0("Plotting ", input$x, " by ", input$y))
  })
}
shinyApp(ui = ui, server = server)
```

### Action Button and Reactive Values

What if we do not want the plot to update right away when we change the inputs? This is particularly useful when there are many inputs, and the output takes a while to load. We can use `actionButton` to control when the plot should be updated.

`reactiveValues` is commonly used to store values, which can be read by other reactive expressions (see above section on `observe` and `reactive`). `observeEvent` is similar to `observe` (see above section) but it only updates when at least one of the event expressions occurs (in the example below, the action button is clicked).

```{r}
ui <- fluidPage(
    titlePanel("My Shiny App"),
  sidebarLayout(
      sidebarPanel(
        selectInput(inputId = 'x', label = "X Variable", choices = names(iris)),
        radioButtons(inputId = 'y', label = "Y Variable", choices = names(iris), inline = FALSE), 
        tags$hr(), 
        actionButton(inputId = 'action', label = "Update Plot")
      ),
      mainPanel(
        plotOutput(outputId = 'plot')
      )
    )
  )
server = function(input, output, session) {
  rv <- reactiveValues(p = NULL)
  
  observe({
    updateRadioButtons(session, 'y', choices = names(iris)[!names(iris) %in% input$x])
  })
  
  dat <- reactive({
    return(data.frame(x = iris[, input$x], y = iris[, input$y]))
  })
  
  output$plot <- renderPlot({
    rv$p
  })
  
  observeEvent(input$action, {
    rv$p <- ggplot(dat()) + 
      geom_point(aes(x = x, y = y, size = 1), color = "#6f22b6") + 
      theme_bw() + 
      theme(legend.position = 'none') +
      labs(x = paste(input$x), y = paste(input$y), title = paste0("Plotting ", input$x, " by ", input$y))
  })
}
shinyApp(ui = ui, server = server)
```

### Useful Resources

- [Shiny Tutorial](https://shiny.rstudio.com/tutorial/): This is a really good place to start. There are so many examples and cool things to be inspired by!
- [Another Tutorial](http://rstudio.github.io/shiny/tutorial/)

### More Advanced, but FUN, Stuff

- [Shiny Widget](https://dreamrs.github.io/shinyWidgets/articles/intro_shinyWidgets_fr.html): Make your Shiny applications prettier with these amazing add-on widgets. My favorite are `actionBttn` and `pickerInput`.
- [Shiny Dashboard](https://rstudio.github.io/shinydashboard/): This package makes organizing your Shiny apps easier than ever.
- [shinyURL](https://github.com/aoles/shinyURL): This lets you share your app easily by including all the selected inputs in the link.
- [plotly](https://plot.ly/r/): Makes `ggplot` interactive (with tooltips)!
- [leaflet](https://rstudio.github.io/leaflet/): Leaflet is an interactive map package. It works extremely well with Shiny. 
- [dygraphs](https://rstudio.github.io/dygraphs/): Interactive time series plots which allows for zooming and panning. 
- and much more!